---
title: "BDM600 - Lab 3 - Group 8"
author: "Ran Arino; Zubeka Dane Dang; Solmaz Heidar Nassab"
date: "2024-02-12"
output: pdf_document
---

## Credentials
All members participated in this lab assignment. The file is the merged version.

## Start
```{r}
# install libraries
library(xtable)
library(vcd)
library(gnm)
library(grid)
library(vcdExtra)
```

## Exercise (1)
```{r}
# get the dataset list
ds <- datasets(package = c("vcd", 'vcdExtra'))
str(ds, vec.len = 2)

# (a): show the number of datasets on each package
table(ds$Package)

# (b): show the tabluer display based on Package and class
table(ds$Package, ds$class)

# (c): examine help files
help(Employment)
example(Employment)
```
- (a): The outpus shows the 78 observations; 33 datasets in vcd and 45 datasets in vcdExtra.
- (b): Shown above
- (c): Retrieved the following description: Data from a 1974 Danish study given by Andersen (1991) on the employees who had been laid off. The workers are classified by their employment status on 1975-01-01, the cause of their layoff and the length of employment before they were laid off.

## Exercise (2)

### (a) Abortion opinion data: Abortion
```{r}
data(Abortion, package="vcdExtra")
str(Abortion)
help(Abortion)
```
* The "Support_Abortion" variable appears to be the response variable, indicating whether individuals support abortion or not.
* The "Sex" and "Status" variables could potentially serve as explanatory variables.
* "Sex" and "Support_Abortion"variables appear to be unordered (nominal) variables as they represent categorical data without inherent ordering. "Status" factor should be treated as ordered of "low" and "hight".
* Substantive questions of interest for analysis of the data might include:

- Is there a significant difference in abortion support between males and females?
- Does socioeconomic status (as represented by the "Status" variable) influence opinions on abortion?

### (b) Caesarian Births: Caesar
```{r}
data(Caesar, package="vcdExtra")
str(Caesar)
help(Caesar)
```
* "Infection" is regarded as the response variable.
* "Risk", "Antibiotics", and "Planned" serve as explanatory variables.
* All four factors appear to be unordered (nominal) variables as they represent categorical data without inherent ordering.
* Substantive questions of interest for analysis of the data:
- What factors are associated with the likelihood of different types of infections during birth?
- Does the use of antibiotics affect the risk of infection?

### (c) Dayton Survey: DaytonSurvey
```{r}
data(DaytonSurvey, package="vcdExtra")
str(DaytonSurvey)
help(DaytonSurvey)
```
* The factors "cigarette", "alcohol", "marijuana", and "Freq" serve as response variables.
* "sex" and "race" typically treated as explanatory variables.
* "Freq" can be treated as orders from low to high number, while others factors are unordered (nominal) variables as they represent categorical data without inherent ordering.
* Substantive questions of interest for analysis of the data:
- Is there a relationship between substance use and demographic characteristics such as sex or race?
- What is the largest category among the three addicted products for those who use them frequently, such as more than 50 or 100 times?

### (d) Minnesota High School Graduates: Hoyt
```{r}
data(Hoyt, package="vcdExtra")
str(Hoyt)
help(Hoyt)
```
* "Status" is natural to consider as the response variable.
* "Rank", father's "Occupation", and "Sex" are explanatory variables.
* "Rank" and "Occupation" can be treated as ordered, while "Status" and "Sex" are unoerders (nominal) variables.
* Substantive questions of interest for analysis of the data:
- Is there a relationship between post-high school status and father's occupation or high school graduate rank?


## Exercise (3)
```{r}
data(UCBAdmissions, package="vcdExtra")
str(UCBAdmissions)
help(UCBAdmissions)
ftable(UCBAdmissions)

# (a): Total number of cases contained in this table
total_cases <- sum(UCBAdmissions)
print(total_cases)

# (b): For each department, find the total number of applicants
total_applicants_by_dept <- apply(UCBAdmissions, 3, sum)
print(total_applicants_by_dept)

# (c): For each department, find the proportion of applicants who were admitted
admitted <- UCBAdmissions["Admitted", , ]
total_admitted_by_dept <- apply(admitted, 2, sum)
proportion_admitted_by_dept <- total_admitted_by_dept / total_applicants_by_dept
print(proportion_admitted_by_dept)

# (d): Construct tabluar with row of department and column of gender.
#  get admitted applicants by gender and department
admitted_by_gender_dept <- apply(UCBAdmissions["Admitted", , , drop = FALSE], c("Dept", "Gender"), sum)
#  get total applicants by gender and department
total_by_gender_dept <- apply(UCBAdmissions, c("Dept", "Gender"), sum)
#  calculate proportions
prop_admitted_by_gender_dept <- admitted_by_gender_dept / total_by_gender_dept
print(prop_admitted_by_gender_dept)
```

## Exercise (4)
```{r}
data(DanishWelfare, package="vcd")
str(DanishWelfare)

# (a): Total number of cases
total_cases <- sum(DanishWelfare$Freq)
print(total_cases)

# (b): Change variables "Alcohol" and "Income" to make them ordered.
DanishWelfare$Alcohol <- ordered(DanishWelfare$Alcohol)
DanishWelfare$Income <- ordered(DanishWelfare$Income)

# (c): Convert to table form
DanishWelfare.tab <- xtabs(Freq ~ Alcohol + Income + Status + Urban, data = DanishWelfare)

# (d): display total frequencies for each category of Urban & Collapse categories.
#  frequency for each category
total_Freq_In_Urban <- aggregate(Freq ~ Urban, data = DanishWelfare, sum)
print(total_Freq_In_Urban)
#  collapse
non_city_list <- c("Copenhagen", "SubCopenhagen", "LargeCity", "Country")
DanishWelfare$Urban <- ifelse(DanishWelfare$Urban %in% non_city_list, "City", "Non-city")
DanishWelfare.tab.collapsed <- xtabs(Freq ~ Alcohol + Income + Status + Urban, data = DanishWelfare)
print(DanishWelfare.tab.collapsed)

# (e): Use structable() or ftable() for a flattened display.
DanishWelfare.ftable <- ftable(DanishWelfare.tab.collapsed)
print(DanishWelfare.ftable)
```

## Exercise (5)
```{r}
data("UKSoccer", package = "vcd")
ftable(UKSoccer)

# (a): Verify the total number of games
games <- apply(UKSoccer, c(1), sum)
print(sum(games))

# (b): Marginal total of goals scored by each of the home and away teams
addmargins(UKSoccer)

# (c): Express each of the marginal totals as proportions
prop.table(UKSoccer)

# (d): Comment on the dist. of the numbers of home-team and away-team goals
marginal_home_goals <- rowSums(UKSoccer)
marginal_away_goals <- colSums(UKSoccer)

# Calculate the mean number of goals for home and away teams
mean_home_goals <- sum(marginal_home_goals * seq_along(marginal_home_goals)) / sum(marginal_home_goals)
mean_away_goals <- sum(marginal_away_goals * seq_along(marginal_away_goals)) / sum(marginal_away_goals)

# Homw: 
print(mean_home_goals)
# Away: 
print(mean_away_goals)
```
From the result above, home teams score more goals on average compared to the scores of the away teams.

## Exercise (6)
```{r}
data("Saxony", package = "vcd")
print(str(Saxony))

data("Geissler", package = "vcdExtra")
print(str(Geissler))

# (a): create data frame
sax12 <- subset(Geissler, Geissler$size == 12)
print(sax12)

# (b): Select the columns for boys and Freq.
sax12_boy <- subset(sax12, select = c(boys, Freq))
print(sax12_boy)

# (c): Use xtabs() with a formula, Freq ~ boys, to create the one-way table
oneway_sax12_boy <- xtabs(Freq ~ boys, data = sax12_boy)
print(oneway_sax12_boy)

# (d): same step for size == 11.
Saxony11 <- subset(Geissler, Geissler$size == 11)
oneway_saxony11_boy <- subset(Saxony11, select = c(boys, Freq))
print(oneway_sax12_boy)
```

## Exercise (7)
```{r}
# (a)
# load data; might need to execute this code
#require(graphics)
# convert UCBAdmissions to data frame
ucb_df <- as.data.frame(UCBAdmissions)
# maipulate factors with interaction
ucb_df$AdmitGender <- with(ucb_df, interaction(Admit, Gender))
# convert back to table
UCB.tab2_long<- xtabs(Freq ~ AdmitGender + Dept, data = ucb_df)
# display table
structable(UCB.tab2_long)
```

```{r}
# (b)
# Use ftable
ucb_ftable <- ftable(Admit ~ Gender + Dept, data = UCBAdmissions)
# convert ftable to a matrix
UCB.tab2_short <- as.matrix(ucb_ftable)
UCB.tab2_short

# Use structable 
ucb_structable <- structable(Admit ~ Gender + Dept, data = UCBAdmissions)
UCB.tab2_structable <- as.matrix(ucb_structable)
UCB.tab2_structable
```

## Exercise (8)
```{r}
# show data info
data("VisualAcuity", package = "vcd")
str(VisualAcuity)

# (a): 4 x 4 freq table for each gender
xtabs(VisualAcuity)

# (b): organized tabular display
structable(VisualAcuity)

# (c): HTML table
xtable(VisualAcuity, type = "html")
```

