---
title: "BDM600 - Lab 7 - Group 8"
author: "Ran Arino; Zubeka Dane Dang; Solmaz Heidar Nassab"
date: "2024-04-03"
output: pdf_document
---

## Credentials
All members participated in this lab assignment. The file is the merged version.

## Start
```{r}
library(vcd)
library(vcdExtra)
library(logmult)
library(MASS)
library(Lahman)
```

## 5.1
```{r}
# load data
data("criminal", package = "logmult")
criminal
```

### (a)
```{r}
model <- loglm(~ Year + Age, data = criminal)
summary(model)
```
- Both the likelihood ratio and Pearson chi-squared tests have very low p-values (much less than 0.05).
- This result supports that the dropping of charges is associated with age, and this relationship has charged over the years recorded in the dataset.

### (b)
```{r}
# mosaic with the option shade = TRUE
mosaic(criminal, shade = TRUE, legend = FALSE)
# Friendly shade
mosaic(criminal, shade = TRUE, legend = FALSE, gp = shading_Friendly)
```
- First plot: Darker shades in the combination of a few groups indicate high frequencies, Year 1995 and Age 19, as well as Year 1958 and Age 16.
- Second plot: If we see that the cells with solid lines (which indicate higher-than-expected frequencies of dropped charges) are more prevalent in the younger age groups as time progresses from 1955 to 1958.
- To wrap up, the mosaic plots, especially with Friendly shade, shows that there is an observable trend where younger men aged 15-19 were more likely to have charges dropped as time went on from 1955 to 1958.

## 5.2
```{r}
data("AirCrash", package = "vcdExtra")
aircrash_tab <- xtabs(~ Phase + Cause, data = AirCrash)
aircrash_tab
```

### (a)
```{r}
mosaic(aircrash_tab, shade = TRUE, legend = FALSE)
```
- Here are the combinations that show higher-than-expected frequencies.
  - En route vs criminal
  - Landing vs Weather
  - Take-off vs mechanical
- However, interpreting the cells labeled as "unknown" may pose challenges due to insufficient information.

### (b)
```{r}
# shortest possible labels
abbrev <- list(abbreviate = TRUE, rot_labels = TRUE, cex = 0.8)
mosaic(aircrash_tab, shade = TRUE, legend = FALSE, labeling_args = abbrev)

# alternate labels
mosaic(aircrash_tab, shade = TRUE, alternate_labels = TRUE, legend = FALSE)

```

### (c)
```{r}
# Calculate marginal frequencies for each factor
phase_freq <- sort(margin.table(aircrash_tab, 1), decreasing = TRUE)
cause_freq <- sort(margin.table(aircrash_tab, 2), decreasing = TRUE)

# Reorder factors in the data frame
AirCrash$Phase <- factor(AirCrash$Phase, levels = names(phase_freq))
AirCrash$Cause <- factor(AirCrash$Cause, levels = names(cause_freq))

# Now create the mosaic plot with the reordered factors
abbrev <- list(abbreviate = TRUE, rot_labels = TRUE, cex = 0.8)
mosaic(~ Phase + Cause, data = AirCrash, shade = TRUE, legend = FALSE, labeling_args = abbrev)

```

## 5.3
```{r}
# renamed the file from "Master" to "People"
data("People", package = "Lahman")
basehands <- with(People, table(throws, bats))

mosaic(basehands, shade = TRUE, legend = FALSE)
```
- There are more than expected number of people who throw and swing with the same handedness (left-handed throwers typically bat left-handedly, and right-handed throwers predominantly bat right-handedly).
- On the other hand, there are less than expected number of people who throw and swing with different handedness. 

## 5.7
```{r}
# load data
data("caith", package = "MASS")
caith
```

### (a)
```{r}
# create matrix
caith.mat <- as.matrix(caith)
dimnames(caith.mat)
names(dimnames(caith.mat)) <- c("Eye", "Hair")
caith.mat
```

### (b)
```{r}
# fit the model using loglm()
caith.model <- loglm(~ Hair + Eye, data = caith.mat, fitted = TRUE)
caith.model
```

### (c)
```{r}
# calculate residuals for this model
residuals(caith.model)
```

### (d)
```{r}
# display mosaic
mosaic(caith.mat, shade = TRUE, legend = FALSE)
```