---
title: "BDM600 - Lab 8 - Group 8"
author: "Ran Arino; Zubeka Dane Dang; Solmaz Heidar Nassab"
date: "2024-04-03"
output: pdf_document
---

## Credentials
All members participated in this lab assignment. The file is the merged version.

## Start
```{r}
library(vcd)
library(vcdExtra)
library(logmult)
library(MASS)
library(ca)
```

### 6.1
#### (a)
```{r}
# load data
data(JobSat)
# simple correspondence analysis
JobSat.ca <- ca(JobSat)
summary(JobSat.ca, rows = FALSE, columns = FALSE)
```
- The inertia accounted for by the one-dimensional solution is 76.4%, which indicates that the first dimension captures a significant portion of the variability in the data.
- The inertia in two-dimensional solution shows an additional 19.7%, which means the the culumative percentage captures 96.1% of variability in the data.

#### (b)
```{r}
# 2D CA solutiton
plot(JobSat.ca)
```
- Point "> 40k" shows the longest vector and being close to the x-axis, which shows the most significant impact on the data distribution, its variance captured by the first dimension.
- Points "<15K", "15-25K", and "ModerateS" have the same angle of vectors, which means that they have similar pattern of association with other categories, although their impact on variance varies.
- Points "VeryD" and "LittleD" close to the y-axis, which shows they are captured by the second dimension.
- Points "<15k" and "ModerateS" are close to each other, which means that the nature of their data are similar.
- Overall, the first dimension can handle the satisfaction levels (worse in left or better in right side) and the income level (lower&medium in left side vs higher incomes in right side)

### 6.2
```{r}
data(criminal)
criminal
```
#### (a)
```{r}
# display ca summary
criminal.ca <- ca(criminal)
summary(criminal.ca, rows = FALSE, columns = FALSE)
```
- The result shows that the accounted for by the one-dimensional solution is more than 90%.
- Hence, almost all portions of the variability are captured in a single dimension.

#### (b)
```{r}
# plot ca
plot(criminal.ca)
```
- Similarity is identified among points "15", "16", and "1958"; they have the same direction and length of vector.
- The category points for both year and age vary systematically over Dimension 1. 
- There were more younger men in later years, and more older in earlier years.

### 6.3
```{r}
# load data
data(AirCrash)
# form a table
AirCrash.tab <- xtabs(~ Phase + Cause,  data = AirCrash)
AirCrash.tab
```

#### (a)
```{r}
# show ca summary
AirCrach.ca <- ca(AirCrash.tab)
summary(AirCrach.ca, rows = FALSE, columns = FALSE)
```
- The 65.6% of variance is accounted for the first dimension.
- The 27.5% of variance is accounted for the second dimension; the 93.1% of data can be explained in tatal of the first and second dimensions.

#### (b)
```{r}
# plot ca
plot(AirCrach.ca)
```
- Overall, "mechanical" and "human-related or weather" causes are more associated with specific flight phases (take-off and landing), respectively.
- Both points of "standing" and "criminal" are relatively rare cases, compared to othe categories.
- Dimension 1 primarily differentiates incidents occurring during standing and landing phases, with a focus on human error and weather problems.
- Dimension 2 primarily differentiates incidents occurring during standing and take-off phases, with a greater emphasis on mechanical problems during take-off.
  
#### (c)
```{r}
# plot ca; map 'symbiplot'
plot(AirCrach.ca, map = "symbiplot", arrows = c(FALSE, TRUE))
```

### 6.4
```{r}
# load data
data(caith)
caith
```

#### (a)
```{r}
# show ca summary
caith.ca <- ca(caith)
summary(caith.ca, rows=FALSE, columns=FALSE)
```
- The analysis shows that associations can be explained by a one-dimensional answer, which is accounted by 86.6%. Adding a second dimension solution essentially complete, which achieved 99%. 
- Additional third dimension does not add much to the explanation of how the variables are connected.

#### (b)
```{r}
# display ca plot
plot(caith.ca, lines = TRUE)
```
- We can observe that while Dimension 1 represents the primary light-dark dimension for hair and eye color, accounting for nearly 86.6% of the association between hair and eye color corresponds to dark (left) vs. light (right) on both variables.
- Dimension 2 could be called "extremes vs. middle", but in CA results with largely 1D association.